---
title: "Bombus population genomics"
output: correlate_gene_density_with_nuc_diversity.html
author: Joe Colgan, Yannick Wurm http://wurmlab.com
---

# Introduction:  
This script investigates the relationship between SNP and gene density.
This script takes in an tab-delimited input file containing six columns:
1) Chromosome name for where a 100kb sliding window analysis is performed
2) Genomic position of the start (bp) of a 100kb sliding window
3) Genomic position of the end (bp) of a 100kb sliding window 
4) Nucleotide diversity for a 100kb sliding window as calculated by [PopGenome](https://cran.r-project.org/web/packages/PopGenome/index.html).
5) Gene density within a 100kb sliding window.
6) Number of intronic SNPs within a 100kb sliding window.

The script performs a correlation test between:
1) Gene density and nucleotide diversity
2) Gene density and intronic SNP density
The script also generates a scatterplot to visualise the potential relationsip between SNP and gene density.

1. Load libraries:

```{r, message = FALSE}
## Load libraries:
libraries <- c("lintr",
               "ggplot2",
               "ggpubr")
for (lib in libraries) {
        if (require(package = lib, character.only = TRUE)) {
                print("Successful")
        } else {
                print("Installing")
                source("https://bioconductor.org/biocLite.R")
                biocLite(pkgs = lib)
                library(lib, character.only = TRUE )
        }
}

## Create output folder:
dir.create(path = "./results/figures",
           recursive = TRUE)
```

2. Read in input file:

```{r, message = FALSE}
data <- read.table(file = "./results/input_for_r.txt",
                   col.names = c("chrom",
                                 "start",
                                 "end",
                                 "nuc_div",
                                 "gene_density",
                                 "intronic_snps"))

## Subset region of interest:
region_of_interest <- subset(data,
                             chrom == "NC_015762.1" &
                                     start > 250000 &
                                     end < 650000)

## Test for correlation between:
## 1) Gene density and nucleotide diversity:
cor.test(x = data$gene_density,
         y = data$nuc_div,
         method = "pearson")

## 2) Gene density and intronic SNPs:
cor.test(x = data$gene_density,
         y = data$intronic_snps,
         method = "pearson")
```         

3. Generate plots:

```{r, message = FALSE)
## Plot:
gene_density <- ggplot(data = data,
       aes(x = gene_density,
           y = nuc_div)) +
        geom_point(color = "light blue",
                   alpha = 0.5) +
        geom_point(data = region_of_interest,
                   aes(x = gene_density,
                       y = nuc_div),
                   color = "black",
                   alpha = 0.8)  +
        ylab("Nucleotide diversity (Pi/100kb)") +
        xlab("Gene Density (Genes/100kb)") +
        theme_bw() +
        theme(axis.title = element_text(size = 15,
                                        face = "bold"),
              axis.text = element_text(size = 10,
                                       face = "plain"))

## Add correlation coeffiecient to plot:
gene_density <- gene_density + stat_cor(method = "pearson",
                        label.x = 15,
                        label.y = 0.0035)

intron_density <- ggplot(data = data,
       aes(x = gene_density,
           y = intronic_snps)) +
        geom_point(color = "light blue",
                   alpha = 0.5) +
        geom_point(data = region_of_interest,
                   aes(x = gene_density,
                       y = intronic_snps),
                   color = "black",
                   alpha = 0.8)  +
        ylab("Intronic SNP Density (SNPs/100kb)") +
        xlab("Gene Density (Genes/100kb)") +
        theme_bw() +
        theme(axis.title = element_text(size = 15,
                                        face = "bold"),
              axis.text = element_text(size = 10,
                                       face = "plain"))

## Add correlation coeffiecient to plot:
intron_density <- intron_density + stat_cor(method = "pearson",
                        label.x = 15,
                        label.y = 1000)
```

4. Run lintr:

```{r, message = FALSE}
lintr::lint(file = "correlate_gene_density_with_nuc_diversity.Rmd")
```
