This script takes a tab-delimited file as input containing the following columns:
1) Chromosome name
2) Start position of chromosome
3) End position of chromosome
4) Gene name
5) Strand
6) Status - if you want to highlight a gene of interest

N.B. I tested out the code using the example_data that comes with the package and everything worked. Created my own dataframe using the same column headers in R but there was an issue with the plots making all genes
the same size in the plot. Unsure of what the problem is. Easiest fix
was to output the example data file and manually change values for plots. 
No issues since. 

```{r}
## Load libraries:
library(ggplot2)
library(gggenes)
## Read in file for plotting:
data <- read.table(file = "./input/input_for_plot.txt", header = TRUE)
colnames(data)

## Update levels so plot and legend order match:
levels(data$chromosome) <- c("NT_177515.1",
                             "NC_015771.1",
                             "NW_014567515.1",
                             "NW_020219542.1",
                             "NW_011967025.1",
                             "NW_012026812.1")
levels(data$gene) <- c("Bombus_terrestris_LOC105666162",
                             "Bombus_impatiens_LOC105681507",
                             "Vollenhovia_emeryi_LOC105558474",
                             "Wasmannia_auropunctata_LOC105448935",
                             "Dinoponera_quadriceps_LOC106751470",
                             "Temnothorax_curvispinosus_LOC112459356")

## Generate plot:
test <- ggplot(data, aes(xmin = start,
                         xmax = end,
                         y = chromosome)) +
        facet_wrap(~ chromosome,
                   scales = "free",
                   ncol = 1) +
        geom_gene_arrow(arrowhead_height = unit(3, "mm"),
                        arrowhead_width = unit(1, "mm")) +
        geom_subgene_arrow(data = data,
                           aes(xmin = start,
                               xmax = end,
                               y = chromosome,
                               fill = gene,
                               xsubmin = start_wol_match,
                               xsubmax = end_wol_match),
                           color = "black",
                           alpha = .7) +
        theme_genes()

## Save image:
dir.create(path = "./results")
ggsave(file = "./results/gene_plot.png",
       height = 10,
       width = 8)
```

Just plot Bombus - highlighting regions of similarity to Wolbachia and other insects

```{r, message = FALSE}
## Read in file for plotting:
data <- read.table(file = "./input/input_for_plot_domains.txt", header = TRUE)
colnames(data)

protein_domain <- ggplot(data, aes(xmin = start,
                         xmax = end,
                         y = protein)) +
        facet_wrap(~ protein,
                   scales = "free",
                   ncol = 1) +
        geom_gene_arrow(arrowhead_height = unit(3, "mm"),
                        arrowhead_width = unit(1, "mm")) +
        geom_subgene_arrow(data = data,
                           aes(xmin = start,
                               xmax = end,
                               y = protein,
                               fill = domain,
                               xsubmin = start_domain,
                               xsubmax = end_domain),
                           color = "black",
                           alpha = .7) +
        theme_genes() +
        theme(axis.ticks = element_blank(),
              axis.title.y = element_blank(),
              axis.text = element_blank(),
              legend.title = element_text(face = "bold"),
              legend.position = "top") +
        scale_x_continuous(position = "top")

## Update colours for plotting:
protein_domain <- protein_domain +
        scale_fill_manual(name = "Functional domains",
                          values = c("yellow",
                                     "blue"))
## Save image as object:
save(protein_domain,
     file = "./results/protein_domain.Rdata")
```

Add piecharts:

```{r, message = FALSE}
## Create a blank theme:
blank_theme <- theme_minimal() +
  theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.ticks = element_blank(),
  plot.title = element_text(size = 14,
                            face = "bold")
  )

## Read in files containing:
## Group information and number of protein matches per group:
sequence_1_info <- read.table(file = "./input/group_counts/sequence_1_prot_matches_by_group.txt",
                              header = TRUE)

sequence_1_info$group <- gsub(pattern = "_",
                              replacement = " ",
                              sequence_1_info$group)

sequence_1_piechart <- ggplot(sequence_1_info,
            aes(x = "",
                y = number,
                fill = group)) +
        geom_bar(width = 1,
                 stat = "identity")

sequence_1_piechart <- sequence_1_piechart +
        coord_polar("y",
                    start = 0) +
        blank_theme +
        theme(axis.text.x = element_blank()) +
        scale_fill_brewer(name = "Taxonomic group",
                          palette = "Blues") +
  theme(axis.title = element_blank(),
        legend.title = element_text(face = "bold"))

## read in sequence 2:
sequence_2_info <- read.table(file = "./input/group_counts/sequence_2_prot_matches_by_group.txt",
                              header = TRUE)

sequence_2_info$group <- gsub(pattern = "_",
                              replacement = " ",
                              sequence_2_info$group)

sequence_2_piechart <- ggplot(sequence_2_info,
            aes(x = "",
                y = number,
                fill = group)) +
        geom_bar(width = 1,
                 stat = "identity")

sequence_2_piechart <- sequence_2_piechart +
        coord_polar("y",
                    start = 0) +
        blank_theme +
        theme(axis.text.x = element_blank()) +
        scale_fill_brewer(name = "Taxonomic group",
                          palette = "Blues") +
        theme(axis.title = element_blank(),
              legend.title = element_text(face = "bold"))

## read in sequence 3:
sequence_3_info <- read.table(file = "./input/group_counts/sequence_3_prot_matches_by_group.txt",
                              header = TRUE)

sequence_3_info$group <- gsub(pattern = "_",
                              replacement = " ",
                              sequence_3_info$group)

sequence_3_piechart <- ggplot(sequence_3_info,
            aes(x = "",
                y = number,
                fill = group)) +
        geom_bar(width = 1,
                 stat = "identity")

sequence_3_piechart <- sequence_3_piechart +
        coord_polar("y",
                    start = 0) +
        blank_theme +
        theme(axis.text.x = element_blank()) +
        scale_fill_brewer(name = "Taxonomic group",
                          palette = "Blues") +
        theme(axis.title = element_blank(),
              legend.title = element_text(face = "bold"))

## read in sequence 4:
sequence_4_info <- read.table(file = "./input/group_counts/sequence_4_prot_matches_by_group.txt",
                              header = TRUE)

sequence_4_piechart <- ggplot(sequence_4_info,
                              aes(x = "",
                                  y = number,
                                  fill = group)) +
        geom_bar(width = 1,
                 stat = "identity")

sequence_4_piechart <- sequence_4_piechart +
        coord_polar("y",
                    start = 0) +
        blank_theme +
        theme(axis.text.x = element_blank()) +
        scale_fill_brewer(name = "Taxonomic group",
                          palette = "Blues") +
        theme(axis.title = element_blank(),
              legend.title = element_text(face = "bold"))
```

Combine plots:

```{r, message = FALSE}
require(ggpubr)
combined_pies <- ggarrange(sequence_1_piechart,
          sequence_2_piechart,
          sequence_3_piechart,
          sequence_4_piechart,
          nrow = 1,
          ncol = 4,
          common.legend = TRUE,
          legend = "bottom")

ggarrange(protein_domain,
          combined_pies,
          ncol = 1,
          nrow = 2,
          align = "hv")

ggsave(file = "./results/protein_domain.png",
       height = 6,
       width = 10)
```

Run lintr:
```{r, message = FALSE}
require(lintr)
lintr::lint(file = "./make_gggenes_plot.Rmd")
```


```{r, message = FALSE}
lintr::lint(file = "./make_gggenes_plot.Rmd")
```