---
title: "Bombus population genomics"
output: plot_wolbachia_sweep.html
author: Joe Colgan, Yannick Wurm http://wurmlab.com
---


## Load libraries:

```{r, message = FALSE}
## Load libraries:
libraries <- c("devtools",
               "ggplot2",
               "ggmap",
               "SNPRelate",
               "reshape",
               "ggpubr",
               "gplots",
               "stringr",
               "gggenes")
for (lib in libraries) {
        if (require(package = lib, character.only = TRUE)) {
                print("Successful")
        } else {
                print("Installing")
                source("https://bioconductor.org/biocLite.R")
                biocLite(pkgs = lib)
                library(lib, character.only = TRUE )
        }
}
```

This script plots nsl score and nucleotide diversity estimates. It also loads an arrow-plot showing the 
distribution of gene locations in the region of interest. It also loads a SNP heatmap showing the presence of shared haplotypes amongst individuals.

```{r, message = FALSE}
## Read in data:
nsl_input <- "./input/nsl_output/NC_015771.1.nsl.out.100bins.norm.abs.bed.tmp"
nsl_data <- read.table(file = nsl_input,
                       col.names = c("chrom",
                                     "position",
                                     "position",
                                     "nsl_score"))
## Both nucleotide diversity and Tajima D are in the same input file:
nuc_div_input <- "./input/nucleotide_diversity_output/bter_all_chrom_10kb.txt"
nuc_div_data <- read.table(file = nuc_div_input,
                           header = TRUE)
```

The gene of interest is an uncharacterised gene (LOC105666162), which codes for proteins with sequence similarity to proteins coded for by Wolbachia-like bacteria. The gene is located on chromosome 11:
NC_015771.1 (1755259..1765481)

```{r, message=FALSE}
## Assign the coordinates of the gene to variables plus 100kb 
## up and downstream of the gene:
chrom <- "NC_015771.1"
start <- 1749721 - 100000
end <- 1776527 + 100000
```

Subset the regions of interest:

```{r, message=FALSE}
## Subset nsl values for plotting:
nsl_input_subset <- subset(nsl_data,
                            position > start &
                            position < end)

## subset sites above 2.56 to highlight on plot:
nsl_snps_highlight <- subset(nsl_input_subset,
                            nsl_score > 2.56)
```

Generate plot of EHH:

```{r, message = FALSE}
## Plot nsl values:
nsl_plot <- ggplot(nsl_input_subset,
                   aes(x = position,
                       y = nsl_score)) +
                xlab("Genomic coordinates (bp)") +
                ylab("|nsl| value") +
                geom_rect(aes(xmin = 1749721,
                              xmax = 1776527,
                              ymin = -Inf,
                              ymax = Inf),
                          fill = "light blue",
                          alpha = 0.008) +
        geom_hline(yintercept = c(2.56),
                   linetype = "dashed",
                   colour = "orange") +
        scale_x_continuous(labels = scales::comma,
                           limits = c(start, end)) +
                geom_point(colour = "black",
                           size = 3,
                           alpha = 0.5) +
                geom_point(data = nsl_snps_highlight,
                           aes(x = position,
                               y = nsl_score),
                           colour = "blue",
                           size = 3,
                           alpha = 0.5) +
                theme_bw() +
                theme(axis.title.x = element_blank(),
                      axis.text.x = element_blank(),
                      axis.ticks.x = element_blank(),
                      axis.title.y = element_text(size = 12,
                                                 face = "bold",
                                                 colour = "black")) +
        theme(plot.margin = unit(c(1, 1, 0, 1), "cm"))

nsl_plot <- nsl_plot +
        theme(axis.text = element_text(size = 12),
              axis.title = element_text(size = 12,
                                        face = "bold"))
```

Subset genomic region to plot nucleotide diversity:

```{r, message = FALSE}
## Subset genomic region of interest:
nuc_div_subset <- subset(nuc_div_data,
                         chrom == "NC_015771.1" &
                         start > 1655259 &
                         end < 1865481)

## Calculate mean nucleotide diversity:
nucdiv_median <- median(nuc_div_data$nuc_diversity)


nuc_div_plot <- ggplot(nuc_div_subset,
                       aes(x = midpoint,
                           y = nuc_diversity)) +
        geom_point(colour = "black",
                   size = 4,
                   alpha = 0.75) +
        xlab("Genomic coordinates (bp)") +
        ylab("Nucleotide diversity") +
        geom_vline(xintercept = c(start + 100000,
                                  end - 100000),
                   linetype = "dashed",
                   colour = "blue") +
        geom_hline(yintercept = c(nucdiv_median),
                   linetype = "dashed",
                   colour = "orange") +
                geom_rect(aes(xmin = 1749721,
                              xmax = 1776527,
                      ymin = -Inf,
                      ymax = Inf),
                  fill = "light blue",
                  alpha = 0.008) +
        scale_x_continuous(labels = scales::comma,
                           limits = c(start, end)) +
        theme_bw()


nuc_div_plot <- nuc_div_plot +
        theme(axis.text = element_text(size = 12),
              axis.title = element_text(size = 12,
                                        face = "bold"))
```

Read in SNP genotype matrix:

```{r, message=FALSE}
## Load previously generated arrow and heatmap plots:
load(file = "results/plots/wolbachia_arrow_plot.Rdata")
load(file = "results/plots/wolbachia_snp_plot.Rdata")
```

Generate combined plots:

```{r, message=FALSE}
## Generate a combined plot:
ggarrange(arrow_plot,
          nsl_plot,
          nuc_div_plot,
          heatmap_plot,
          nrow = 4,
          ncol = 1,
          labels = c("A",
                     "B",
                     "C",
                     "D"),
          heights = c(1,
                     1.8,
                     2,
                     3),
          align = "hv")

## Save:
ggsave(file = "results/wolbachia_combined_plot.png",
       width = 10,
       height = 10)
```

Run lintr:

```{r, message = FALSE}
## Run lintr:
lintr::lint(file = "./plot_wolbachia_sweeps_attic.Rmd")
```