---
title: "Bombus population genomics"
output: plot_gggenes_arrows_chrom_1.html
author: Joe Colgan, Yannick Wurm http://wurmlab.com
---

## Introduction
This script takes a tab-delimited file as input containing the following columns:
1) Chromosome name
2) Start position of chromosome
3) End position of chromosome
4) Gene name
5) Strand
6) Status - if you want to highlight a gene of interest

N.B. I tested out the code using the example_data that comes with the package and everything worked. Created my own dataframe using the same column headers in R but there was an issue with the plots making all genes
the same size in the plot. Unsure of what the problem is. Easiest fix
was to output the example data file and manually change values for plots. 
No issues since. 

```{r, message = FALSE}
## Load libraries:
libraries <- c("ggplot2",
               "ggpubr",
               "gggenes",
               "lintr")
for (lib in libraries) {
        if (require(package = lib, character.only = TRUE)) {
                print("Successful")
        } else {
                print("Installing")
                source("https://bioconductor.org/biocLite.R")
                biocLite(pkgs = lib)
                library(lib, character.only = TRUE )
        }
}

dir.create(path = "./results/plots",
           recursive = TRUE)
```

```{r, message = FALSE}
## Read in file for plotting:
input <- "./populations/bedfiles/region_two_sites_for_gggenes_input.txt"
data <- read.table(file = input,
                   col.names = c("molecule",
                                "start",
                                "end",
                                "gene",
                                "strand"))

## Assign values to start and end positions:
start_position <- data$start[1]
end_position <- data$end[nrow(data)]

data$status <- data$start >= 25032154 & data$end <= 25143386

## Turn off scientific notation:
options(scipen = 999)

data$direction <- ifelse(data$strand == "forward", 1, -1)

## Plot:
arrow_plot_two <- ggplot(data, aes(xmin = as.numeric(start),
                              xmax = as.numeric(end),
                              y = strand,
                              fill = status,
                              label = gene,
                              forward = direction)) +
        xlim(start_position, end_position) +
  geom_gene_arrow(arrowhead_height = unit(3, "mm"),
                  arrowhead_width = unit(1, "mm")) +
  geom_gene_label(align = "left",
                  min.size = 6) +
  facet_wrap(~ molecule,
             scales = "free",
             nrow = 1,
             shrink = FALSE) +
  scale_fill_brewer(palette = "Blues") +
        theme_genes() +
        theme(axis.title.y = element_blank(),
              axis.text.y = element_blank(),
              axis.title.x = element_blank(),
              axis.text.x = element_blank(),
              axis.ticks.x = element_blank(),
              legend.position = "none")

## Combine plot:
save(arrow_plot_two,
       file = "./results/plots/chrom_1_arrow_plot_region_2.Rdata")
```

Run lintr

```{r, message = FALSE}
## Run lintr:
lintr::lint(file = "./plot_gggene_arrows_chrom_1_region_2.Rmd")
```

