---
title: "Bombus population genomics"
output: plot_tackykinin_sweep.html
author: Joe Colgan, Yannick Wurm http://wurmlab.com
---


```{r, message = FALSE}
## Load libraries:
libraries <- c("ggplot2",
               "SNPRelate", 
               "reshape2",
               "ggpubr",
               "stringr",
               "gggenes")
for (lib in libraries) {
        if (require(package = lib, character.only = TRUE)) {
                print("Successful")
        } else {
                print("Installing")
                source("https://bioconductor.org/biocLite.R")
                biocLite(pkgs = lib)
                library(lib, character.only = TRUE )
        }
}
```

This script plots nsl score, ihs score, population-scaled recombination rate, nucleotide diversity and Tajima's D estimates.

```{r, message = FALSE}
## Read in data:
nsl_data <- read.table(file = "./input/nsl_output/NC_015762.1.nsl.out.100bins.norm.abs.bed.tmp",
                       col.names = c("chrom",
                                     "position",
                                     "position",
                                     "nsl_score"))
ihs_data <- read.table(file = "./input/ihs_output/NC_015762.1.ihs.out.100bins.norm.abs.bed.tmp",
                       col.names = c("chrom",
                                     "position",
                                     "position",
                                     "ihs_score"))
recomb_data <- read.table(file = "./input/recombination_rate/NC_015762.1.recode.vcf.ldhelmet.ldhelmet.post.txt.tmp.genetic_map.txt.recomb.txt",
                          header = TRUE)
## Both nucleotide diversity and Tajima D are in the same input file:
nuc_div_data <- read.table(file = "./input/nucleotide_diversity_output/bter_all_chrom_10kb.txt",
                           header = TRUE)
```

This region of interest contains 53 genes.

```{r, message=FALSE}
## Assign the coordinates of the gene to variables plus 100kb 
## up and downstream of the gene:
chrom <- "NC_015762.1"
start <- 362267 - 100000
end <- 560366 + 100000
```

Subset the regions of interest:

```{r, message=FALSE}
## Subset nsl values for plotting:
nsl_input_subset <- subset(nsl_data,
                            position > start &
                            position < end)

## subset sites above 2.56 to highlight on plot:
nsl_snps_highlight <- subset(nsl_input_subset, 
                            nsl_score > 2.56)

## Subset ihs values for plotting:
ihs_input_subset <- subset(ihs_data,
                            position > start &
                            position < end)

ihs_snps_highlight <- subset(ihs_input_subset, 
                            ihs_score > 2.89)
```

Generate plot of EHH:

```{r, message = FALSE}
nsl_plot <- ggplot(nsl_input_subset,
                   aes(x = position,
                       y = nsl_score)) +
                xlab("Genomic coordinates (bp)") +
                ylab("|nsl| value") +
                geom_rect(aes(xmin = 362267,
                              xmax = 560366,
                              ymin = -Inf,
                              ymax = Inf),
                          fill = "light blue",
                          alpha = 0.008) +
        geom_hline(yintercept = c(2.56),
                   linetype = "dashed",
                   colour = "orange") +
        scale_x_continuous(labels = scales::comma,
                           limits = c(start, end)) +
                geom_point(colour = "black",
                           size = 3,
                           alpha = 0.5) +
                geom_point(data = nsl_snps_highlight, 
                           aes(x = position,
                               y = nsl_score),
                           colour = "blue",
                           size = 3,
                           alpha = 0.5) +
                theme_bw() +
                theme(axis.title.x = element_blank(),
                      axis.text.x = element_blank(),
                      axis.ticks.x = element_blank(),
                      axis.title.y = element_text(size = 15,
                                                 face = "bold",
                                                 colour = "black")) +
        theme(plot.margin = unit(c(1,1,0,1), "cm"))
```

Generate the ihs plot:

```{r, message = FALSE}
ihs_plot <- ggplot(ihs_input_subset,
                   aes(x = position,
                       y = ihs_score)) +
                xlab("Genomic coordinates (bp)") +
                ylab("|ihs| value") +
                geom_rect(aes(xmin = 362267,
                              xmax = 560366,
                              ymin = -Inf,
                              ymax = Inf),
                          fill = "light blue",
                          alpha = 0.008) +
        geom_hline(yintercept = c(2.89),
                   linetype = "dashed",
                   colour = "orange") +
        scale_x_continuous(labels = scales::comma,
                           limits = c(start, end)) +
                geom_point(colour = "black",
                           size = 3,
                           alpha = 0.5) +
                geom_point(data = ihs_snps_highlight, 
                           aes(x = position,
                               y = ihs_score),
                           colour = "blue",
                           size = 3,
                           alpha = 0.5) +
                theme_bw() +
                theme(axis.title.x = element_blank(),
                      axis.text.x = element_blank(),
                      axis.ticks.x = element_blank(),
                      axis.title.y = element_text(size = 15,
                                                 face = "bold",
                                                 colour = "black")) +
        theme(plot.margin = unit(c(1,1,0,1), "cm"))
```

Plot recombination rate for the region:

```{r, message = FALSE}
## Subset region of genome to plot:
recomb_data_subset <- subset(recomb_data,
                            start.pos > start &
                            start.pos < end)

recomb_plot <- ggplot(recomb_data_subset,
                      aes(x = start.pos,
                          y = recomb.rate.perbp)) +
               geom_point()
```

Subset genomic region to plot nucleotide diversity:

```{r, message = FALSE}
## 
nuc_div_subset <- subset(nuc_div_data,
                         chrom=="NC_015762.1" &
                         start > 362267 - 100000 &
                         end < 560366 + 100000)

## Calculate mean nucleotide diversity:
nucdiv_median <- median(nuc_div_data$nuc_diversity)


nuc_div_plot <- ggplot(nuc_div_subset,
                       aes(x = midpoint,
                           y = nuc_diversity)) +
        geom_point(colour = "black",
                   size = 4,
                   alpha = 0.75) +
        xlab("Genomic coordinates (bp)") +
        ylab("Nucleotide diversity") +
        geom_vline(xintercept = c(start + 100000,
                                  end - 100000),
                   linetype = "dashed",
                   colour = "blue") +
        geom_hline(yintercept = c(nucdiv_median),
                   linetype = "dashed",
                   colour = "orange") +
                geom_rect(aes(xmin = 362267,
                              xmax = 560366,
                      ymin = -Inf,
                      ymax = Inf),
                  fill = "light blue",
                  alpha = 0.008) +
        scale_x_continuous(labels = scales::comma,
                           limits = c(start, end)) +
        theme_bw() 

nuc_div_plot <- nuc_div_plot +
        theme(axis.text = element_text(size = 15),
              axis.title = element_text(size = 15,
                                        face = "bold"))
```

Load previously generated arrow plot:

```{r, message = FALSE}
## Load arrow plot:
load(file = "./results/plots/chrom_1_arrow_plot.Rdata")
load(file = "./results/plots/chrom_1_snp_plot.Rdata")
```

Generate combined plots:

```{r, message=FALSE}
## Generate a combined plot:
ggarrange(arrow_plot,
          nsl_plot,
          nuc_div_plot,
          heatmap_plot,
          nrow = 4,
          ncol = 1,
          labels = c("A",
                     "B",
                     "C",
                     "D"),
          heights = c(1,
                     1.8,
                     2,
                     3),
          align = "hv")

## Save:
ggsave(file = "results/large_sweep_chrom_one_combined_plot.png",
       width = 10,
       height = 10)

```