---
title: "Bombus population genomics"
output: plot_go_barchart.html
author: Joe Colgan, Yannick Wurm http://wurmlab.com
---

## Introduction:  

This script is for visualisation of the output of gene ontology (GO) analyses using the R package TopGo.
This script creates a barchart of the number of significant genes annotated to a particular GO term, with an associated p-value. This script takes an input of .csv files containing an output table of GO analyses.

```{r, message = FALSE}
# Load libraries; install from scratch if needed
libraries <- c("ggplot2",
               "lattice",
               "ggpubr")
for (lib in libraries) {
    if (require(package = lib, character.only = TRUE)) {
        print("Successful")
    } else {
        print("Installing")
        source("https://bioconductor.org/biocLite.R")
        library(lib, character.only = TRUE)
    }
}

## Create output directory:
dir.create("results/figures/",
           recursive = TRUE)
```

## Step One: Load input files
Read in files for the output of Gene Ontology enrichment analysis.

Load results from diapause samples:

```{r, message = FALSE}
## Read in biological processes GO terms:
bp_terms <- read.table(file = "results/output_50_ks_bh_0.01/BP_sig.tsv",
                       header = TRUE)
bp_terms$category <- "BP"

## Read in molecular function GO terms:
mf_terms <- read.table(file = "results/output_50_ks_bh_0.01/MF_sig.tsv",
                       header = TRUE)
mf_terms$category <- "MF"

## Read in molecular function GO terms:
cc_terms <- read.table(file = "results/output_50_ks_bh_0.01/CC_sig.tsv",
                       header = TRUE)
cc_terms$category <- "CC"
```

For each time-point, combine gene ontolgy terms for plotting:

```{r, message = FALSE}
## Combine GO terms:
combined_terms <- rbind(bp_terms,
                        mf_terms,
                        cc_terms)
combined_terms$category <- factor(combined_terms$category)

## Set order of GO categories to plot
levels(combined_terms$category) <- c("BP", "MF", "CC")

## Log transform p values for plotting:
combined_terms$weight_ks_adjusted <- -log(combined_terms$weight_ks_adjusted)

## For plotting, update GO term names to include total number of annotated terms:
combined_terms$updated_terms <- paste(combined_terms$Term,
                                      " ",
                                      "(",
                                      combined_terms$Annotated,
                                      ")",
                                      sep = "")

## Remove underscore:
combined_terms$updated_terms <- gsub("_", " ", combined_terms$updated_terms)

## Generate plot:
combined_terms_plot <- ggbarplot(combined_terms,
                                          x = "updated_terms",
                                          y = "weight_ks_adjusted",
                                          position = position_dodge(0.1),
                                          fill = "category",
                                          color = NULL,
                                          palette = "jco",
                                          sort.val = "asc",
                                          sort.by.groups = TRUE,
                                          ylab = "-log10(p)",
                                          xlab = "Gene Ontology term",
                                          legend.title = "Gene Ontology",
                                          lab.col = "black",
                                          lab.size = 2,
                                          lab.vjust = 0.5,
                                          lab.hjust = 1,
                                 short.panel.labs = FALSE,
                                          legend = "top",
                                          rotate = TRUE,
                                          ggtheme = theme_minimal())

## Make font bigger and bold:
combined_terms_plot <- combined_terms_plot +
        scale_y_continuous(expand = c(0, 2)) +
        theme(axis.text = element_text(size = 15),
              axis.title.x = element_text(size = 15,
                                          face = "bold"),
              axis.title.y = element_text(size = 15,
                                          face = "bold"),
              axis.text.y = element_text(size = 10,
                                         face = "plain"),
              axis.text.x = element_text(size = 10),
              legend.position = "none") +
        geom_hline(yintercept = 1.301,
                   linetype = "dashed",
                   colour = "black")

## Print to console:
combined_terms_plot

## Save picture:
ggsave(file = "results/figures/combined_go_terms_barchart.png",
       height = 16,
       width = 14)

## Save picture:
ggsave(file = "results/figures/combined_go_terms_barchart.pdf",
       dpi = 600,
       height = 16,
       width = 14)

## Save as R object:
save(combined_terms_plot,
     file = "./results/figures/combined_go_terms_barchart.RData")
```

Run lintr:

```{r, message = FALSE}
## Run lintr
lintr::lint(file = "./plot_go_barchart.Rmd")
```
