---
title: "Bombus population genomics"
output: make_map.html
author: Joe Colgan, Yannick Wurm http://wurmlab.com
---

## Introduction:
The purpose of this script is to generate a map of Great Britain and highlight sampling sites.
The output of the script is an image (.png or .pdf) file.

The first step involves installing and loading libraries:
```{r, message = FALSE}
libraries <- c("ggplot2",
               "ggrepel",
               "ggspatial",
               "sf",
               "rnaturalearth",
               "rnaturalearthdata",
               "rgeos")
for (lib in libraries) {
        if (require(package = lib, character.only = TRUE)) {
                print("Successful")
        } else {
                print("Installing")
                source("https://bioconductor.org/biocLite.R")
                biocLite(pkgs = lib)
                library(lib, character.only = TRUE )
        }
}

## Save to file:
dir.create(path = "./results")
```

Load data for plotting from the world:

```{r, message = FALSE}
## Set theme:
theme_set(theme_bw())

## Load image of the world:
world <- ne_countries(scale = "medium",
                      returnclass = "sf")
world_points <- st_centroid(world)
```

For plotting of collection sites, read in input file containing coordinates:

```{r, message = FALSE}
## Read in data:
data_points <- read.csv(file = "input/input_for_maps.csv",
                        header = TRUE)
```

Generate map:

```{r, message = FALSE}
britain_map <- ggplot(data = world) +
        geom_sf() +
        geom_text(data = world_points,
                  aes(x = "Longitude",
                      y = "Latitude",
                      label = name),
                  color = "darkblue",
                  fontface = "bold",
                  check_overlap = FALSE) +
        annotate(geom = "text",
                 x = 0,
                 y = 55,
                 label = "North Sea",
                 fontface = "italic",
                 color = "grey22",
                 size = 8) +
        geom_point(data = data_points,
                   aes(x = Longitude,
                       y = Latitude,
                       fill = Latitude),
                   shape = 21,
                   size = 5) +
        geom_label_repel(data = data_points,
                         aes(x = Longitude,
                             y = Latitude,
                             label = site_number_for_plots),
                         size = 4) +
        coord_sf(xlim = c(-6, 2),
                 ylim = c(57, 50),
                 expand = FALSE) +
        xlab("Longitude") +
        ylab("Latitude") +
        annotation_scale(location = "bl",
                         width_hint = 0.5) +
        theme(legend.position = "none")

## Update colour:
britain_map <- britain_map +
               theme(axis.text = element_text(size = 10,
                                              face = "plain"),
                     axis.title = element_text(size = 15,
                                               face = "bold"))
```

Save image to file:

```{r, message = FALSE}
## Save to file:
ggsave(file = "./results/map_of_britain.png",
       height = 15,
       width = 10)

## Save image as an R object:
save(britain_map,
     file = "./results/map_of_britain.Rdata")
```

Run lintr:

```{r, message = FALSE}
## Run lintr:
lintr::lint(file = "./make_map.Rmd")
```
