---
title: "Bombus population genomics"
output: interpro_enrichment_nsl.html
author: Joe Colgan, Yannick Wurm http://wurmlab.com
---

## Introduction:
This script takes a tab-delimited text file as input containing four columns:
1. Gene ID (locus)
2. |iHS| score
3. |nSL| score
4. List of interproscan domains, which was obtained from ensembl.
This scripts performs a wilcoxon test to identify interpro functional domains significantly enriched in genes with high |nsl| values.
This script outputs a tab-delimited text file containing two columns:
1. interpro domain accession ID
2. Bonferroni-adjusted p-value
This output file can be used to plot a barchart of enriched interpro domains.

```{r, message = FALSE}
## Load libraries:
libraries <- c("lintr")
for (lib in libraries) {
        if (require(package = lib, character.only = TRUE)) {
                print("Successful")
        } else {
                print("Installing")
                source("https://bioconductor.org/biocLite.R")
                biocLite(pkgs = lib)
                library(lib, character.only = TRUE )
        }
}
```

Read in input files:

```{r}
## Read data from file:
nsl_interpro_database <- read.table("./input/nsl/interproscan_input_plus_nsl_top_scores_per_gene.txt",
                                    col.names = c("locus",
                                                  "nsl",
                                                  "length",
                                                  "interpros"),
                                    fill = TRUE)

## Ensure nsl and ihs scores are numeric:
nsl_interpro_database$nsl <- as.numeric(as.character(nsl_interpro_database$nsl))
# we suppose that all values of 0 mean NO nsl was calculated
nsl_interpro_database_nozero <- na.omit(nsl_interpro_database)
nsl_interpro_database$nsl_corrected <- nsl_interpro_database$nsl/as.numeric(unlist(nsl_interpro_database$length))

## Ensure sorted by nsl:
nsl_interpro_database_nozero <- nsl_interpro_database[order(nsl_interpro_database$nsl_corrected), ]

## Read in a file containing interproscans at least found in 20 genes:
interpros_to_check <- read.table("input/database/interproscan_of_interest_input_for_wilcoxon.n20.txt")$V1

wilcox_pvalues <- rep(NA,
               times = length(interpros_to_check))
names(wilcox_pvalues) <- interpros_to_check

ks_pvalues <- rep(NA,
               times = length(interpros_to_check))
names(ks_pvalues) <- interpros_to_check

## For each interpro compare present in genes with high nsl values against
## genes with low nsl values:
for (current_interpro in interpros_to_check) {
        print(paste("testing", current_interpro))
        nsls_with_interpro <- subset(nsl_interpro_database_nozero,
                                     subset = grepl(pattern = current_interpro,
                                                    x = interpros))$nsl_corrected
        nsls_without_interpro <- subset(nsl_interpro_database_nozero,
                                        subset = !grepl(pattern = current_interpro,
                                                        x = interpros))$nsl_corrected
        wilcox_result <- wilcox.test(x = nsls_with_interpro,
                                     y = nsls_without_interpro)
        ks_result <- ks.test(x = nsls_with_interpro,
                                     y = nsls_without_interpro)
        wilcox_pvalues[current_interpro] <- wilcox_result$p.value
        ks_pvalues[current_interpro] <- ks_result$p.value
}

## Adjust p values for multiple testing:
wilcox_pvalues_adjust <- p.adjust(wilcox_pvalues,
                           method = "bonferroni")
ks_pvalues_adjust <- p.adjust(ks_pvalues,
                           method = "bonferroni")

## Convert to dataframe:
pvalues_adjust_df <- cbind(as.data.frame(wilcox_pvalues_adjust),
                           as.data.frame(ks_pvalues_adjust))
pvalues_adjust_df$ipr_domain <- rownames(pvalues_adjust_df)

## Subset significant terms:
nrow(subset(pvalues_adjust_df,
            wilcox_pvalues_adjust < 0.05 &
                    wilcox_pvalues_adjust > 0))
pvalues_adjust_sig <- subset(pvalues_adjust_df,
                             wilcox_pvalues_adjust < 0.05)

## Explore correlation of values for Wilcoxon and KS tests:
cor.test(pvalues_adjust_df$wilcox_pvalues_adjust,
         pvalues_adjust_df$ks_pvalues_adjust)

pvalues_adjust_sig <- pvalues_adjust_sig[order(pvalues_adjust_sig$wilcox_pvalues_adjust), ]

## Write to output:
write.table(pvalues_adjust_sig,
      file = "./results/signficiant_ipr_domains_length_corrected.txt",
      row.names = FALSE,
      col.names = TRUE,
      sep = "\t",
      quote = FALSE)
```

Run lintr:

```{r, message = FALSE}
## Run lintr
lintr::lint("./interpro_enrichment_nsl.Rmd")
```