---
title: "Bombus population genomics"
output: plot_apis_fst.html
authors: Joe Colgan (joscolgan)
---

## Introduction:
This script takes output from [vcftools](http://vcftools.sourceforge.net/) which calculated pairwise Weir and Cockram's Fst between two of four honeybee (_Apis mellifera_) populations.  
The scripts plots scatterplots for two regions of interest in the honeybee genome.  
The script outputs an image (.png, .jpeg, .pdf) of these plots.  

```{r, message = FALSE}
#Load libraries; install from scratch if needed
libraries <- c("ggplot2",
               "PopGenome",
               "stringr",
               "lintr",
               "reshape2",
               "ggpubr")
for (lib in libraries) {
    if (require(package = lib, character.only = TRUE)) {
        print("Successful")
    } else {
        print("Installing")
        source("https://bioconductor.org/biocLite.R")
        library(lib, character.only = TRUE)
    }
}
```

Read in input file:

```{r, message = FALSE}
input <- read.table(file = "./other_population_comparisons.txt",
                    header = FALSE)
colnames(input) <- c("chrom",
                     "position",
                     "african_vs_asian",
                     "african_vs_eEurope",
                     "african_vs_wEurope",
                     "asia_vs_eEurope",
                     "asia_vs_wEurope",
                     "eEurope_vs_wEurope")

## Subset genomic regions of interest:
region_1 <- subset(input,
                   position > (20537332 - 100000) &
                           position < (20625795 + 100000))
region_1$chrom <- NULL

region_1_melt <- melt(region_1, id.vars = "position")

## plot:
region_1_plot <- ggplot(data = region_1_melt,
       aes(x = position,
           y = value,
           alpha = 0.5,
           color = variable)) +
        geom_point()


## Subset genomic regions of interest:
region_2 <- subset(input,
                   position > (25036596 - 100000) &
                   position < (25143386 + 100000))
## Remove chrom column:
region_2$chrom <- NULL

## Reshape dataframe:
region_2_melt <- melt(region_2,
                      id.vars = "position")

## Generate scatterplot for each comparison:
## a) African and Asian populations:
## ai) Region one:
african_vs_asian_region_1 <- ggplot(data = region_1,
                                    aes(x = position,
                                        y = african_vs_asian,
                                        alpha = 0.5,
                                        color = african_vs_asian)) +
        geom_point() +
        xlab("Genomic position (bp)") +
        ylab("Fst") +
        labs(colour = "Fst") +
        ggtitle("A vs O: Region 1") +
        theme_bw() +
        theme(axis.title = element_text(size = 15,
                                        face = "bold"),
              axis.text = element_text(size = 10,
                                       face = "plain"))

## aii) Region two:
african_vs_asian_region_2 <- ggplot(data = region_2,
                                    aes(x = position,
                                        y = african_vs_asian,
                                        alpha = 0.5,
                                        color = african_vs_asian)) +
        geom_point() +
        xlab("Genomic position (bp)") +
        ylab("Fst") +
        labs(colour = "Fst") +
        ggtitle("A vs O: Region 2") +
        theme_bw() +
        theme(axis.title = element_text(size = 15,
                                        face = "bold"),
              axis.text = element_text(size = 10,
                                       face = "plain"))

## b) African and Eastern European populations:
## bi) Region one:
african_vs_eeurope_region_1 <- ggplot(data = region_1,
                                      aes(x = position,
                                          y = african_vs_eEurope,
                                          alpha = 0.5,
                                          color = african_vs_eEurope)) +
        geom_point() +
        xlab("Genomic position (bp)") +
        ylab("Fst") +
        labs(colour = "Fst") +
        ggtitle("A vs M: Region 1") +
        theme_bw() +
        theme(axis.title = element_text(size = 15,
                                        face = "bold"),
              axis.text = element_text(size = 10,
                                       face = "plain"))

## b) African and Eastern European populations:
## bii) Region two:
african_vs_eeurope_region_2 <- ggplot(data = region_2,
                                      aes(x = position,
                                          y = african_vs_eEurope,
                                          alpha = 0.5,
                                          color = african_vs_eEurope)) +
        geom_point() +
        xlab("Genomic position (bp)") +
        ylab("Fst") +
        labs(colour = "Fst") +
        ggtitle("A vs M: Region 2") +
        theme_bw() +
        theme(axis.title = element_text(size = 15,
                                        face = "bold"),
              axis.text = element_text(size = 10,
                                       face = "plain"))

## c) African and Western European populations:
## ci) Region one:
african_vs_weurope_region_1 <- ggplot(data = region_1,
                                      aes(x = position,
                                          y = african_vs_wEurope,
                                          alpha = 0.5,
                                          color = african_vs_wEurope)) +
        geom_point() +
        xlab("Genomic position (bp)") +
        ylab("Fst") +
        labs(colour = "Fst") +
        ggtitle("A vs C: Region 1") +
        theme_bw() +
        theme(axis.title = element_text(size = 15,
                                        face = "bold"),
              axis.text = element_text(size = 10,
                                       face = "plain"))

## cii) Region two:
african_vs_weurope_region_2 <- ggplot(data = region_2,
                                      aes(x = position,
                                          y = african_vs_wEurope,
                                          alpha = 0.5,
                                          color = african_vs_wEurope)) +
        geom_point() +
        xlab("Genomic position (bp)") +
        ylab("Fst") +
        labs(colour = "Fst") +
        ggtitle("A vs C: Region 2") +
        theme_bw() +
        theme(axis.title = element_text(size = 15,
                                        face = "bold"),
              axis.text = element_text(size = 10,
                                       face = "plain"))
```

Generate a plot for African population and others together:

```{r, message = FALSE}
## Generate a multi image plot:
ggarrange(african_vs_asian_region_1,
          african_vs_eEurope_region_1,
          african_vs_wEurope_region_1,
          african_vs_asian_region_2,
          african_vs_eEurope_region_2,
          african_vs_wEurope_region_2,
          nrow = 2,
          ncol = 3,
          align = "hv",
          common.legend = TRUE)

## Save:
ggsave(file = "./results/plot_fst.png",
       height = 10,
       width = 15)
```

Run through lintr:

```{r, message = FALSE}
lintr::lint(file = "./plot_fst.Rmd")
```
