---
title: "Bombus population genomics"
output: supplemental_1_pca_correlation.html
author: Joe Colgan, Yannick Wurm http://wurmlab.com
---

## Introduction:
This script loads R objects containing:
1) A principal component analysis of sites (highlighted by landuse type)
2) A principal component analysis of sites (highlighted by latitude)
The script investigates correlation between PC2 and latitude.
The script outputs a scatterplot (.png, .jpeg, .pdf) and R object.  

```{r, message = FALSE}
# Load libraries; install from scratch if needed
libraries <- c("ggplot2",
               "ggpubr",
               "ggrepel",
               "lintr")
for (lib in libraries) {
    if (require(package = lib, character.only = TRUE)) {
        print("Successful")
    } else {
        print("Installing")
        source("https://bioconductor.org/biocLite.R")
        library(lib, character.only = TRUE)
    }
}
## Create output directory:
dir.create("results")
```
Next step is to load input files (R objects)

```{r, message = FALSE}
## Load objects:
load(file = "./input/pca_plot/pca_plots.Rdata")
```

1) Plot PCA and Latitude:

```{r, message = FALSE}
## Perform correlation test:
cor.test(tab$EV2,
         tab$latitude,
         methd = "pearson")

## Update latitude plot with points:
lat_correlation <- ggplot(tab,
            aes(x = EV2,
                y = latitude)) +
            xlab(paste("Principal component ",
                       2,
                       " ",
                       "(",
                       round(pc_percent[2],
                             2), "%", ")",
                       sep = "")) +
            ylab("Latitude (°N)") +
        geom_point(size = 5,
                aes(color = latitude,
                    fill = latitude)) +
        geom_label_repel(aes(label = site_id)) +
        theme_minimal() +
            theme(axis.text = element_text(size = 10,
                  face = "plain"),
                  axis.title = element_text(size = 15,
                  face = "bold")) +
            theme(legend.position = "none")

## Annotate plot with results from cor.test
lat_correlation <- lat_correlation +
        annotate(geom = "text",
                           y = 55.5,
                           x = -0.15,
                           label = "r = 0.75",
                 size = 5) +
        annotate(geom = "text",
                           y = 55.2,
                           x = -0.15,
                           label = "p < 1e-08",
                 size = 5)

## Save as .png:
ggsave(file = "./results/supplemental_1_pca_correlation.png",
       height = 8,
       width = 10)

## Save as .pdf
ggsave(file = "./results/supplemental_1_pca_correlation.pdf",
       dpi = 600,
       height = 8,
       width = 10)
```

Save also as an R object:

```{r, message = FALSE}
save.image(file = "./results/supplemental_1_pca_correlation.RData")
```

Run lintr:

```{r, message = FALSE}
lintr::lint(file = "./supplemental_1_pca_correlation.Rmd")
```
